---
title: "CFM"
output: flexdashboard::flex_dashboard
---
```{r}
library(flexdashboard)
library(shiny)
library(dplyr)
library(plotly)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(corrplot)
library(MASS)
library(xgboost)
library(caret)
library(CatEncoders)
library(splitTools)
library(ranger)
```

```{r data}
CFM_data <- read.csv("C:/Users/bssup/Documents/fall22/R/project/Food_mart/data.csv")

transformed_data = CFM_data
#define original categorical labels
label = LabelEncoder.fit(CFM_data$food_category)
transformed_data$food_category = transform(label, CFM_data$food_category)

label = LabelEncoder.fit(CFM_data$food_department)
transformed_data$food_department = transform(label, CFM_data$food_department)

label = LabelEncoder.fit(CFM_data$food_family)
transformed_data$food_family = transform(label, CFM_data$food_family)

label = LabelEncoder.fit(CFM_data$promotion_name)
transformed_data$promotion_name = transform(label, CFM_data$promotion_name)

label = LabelEncoder.fit(CFM_data$marital_status)
transformed_data$marital_status = transform(label, CFM_data$marital_status)

label = LabelEncoder.fit(CFM_data$gender)
transformed_data$gender = transform(label, CFM_data$gender)

label = LabelEncoder.fit(CFM_data$education)
transformed_data$education = transform(label, CFM_data$education)

label = LabelEncoder.fit(CFM_data$member_card)
transformed_data$member_card = transform(label, CFM_data$member_card)

label = LabelEncoder.fit(CFM_data$occupation)
transformed_data$occupation = transform(label, CFM_data$occupation)

label = LabelEncoder.fit(CFM_data$avg..yearly_income)
transformed_data$avg..yearly_income = transform(label, CFM_data$avg..yearly_income)

label = LabelEncoder.fit(CFM_data$sales_country)
transformed_data$sales_country = transform(label, CFM_data$sales_country)

label = LabelEncoder.fit(CFM_data$brand_name)
transformed_data$brand_name = transform(label, CFM_data$brand_name)

label = LabelEncoder.fit(CFM_data$houseowner)
transformed_data$houseowner = transform(label, CFM_data$houseowner)

label = LabelEncoder.fit(CFM_data$store_type)
transformed_data$store_type = transform(label, CFM_data$store_type)

label = LabelEncoder.fit(CFM_data$store_city)
transformed_data$store_city = transform(label, CFM_data$store_city)

label = LabelEncoder.fit(CFM_data$store_state)
transformed_data$store_state = transform(label, CFM_data$store_state)

label = LabelEncoder.fit(CFM_data$media_type)
transformed_data$media_type = transform(label, CFM_data$media_type)


#split the data 


# Split data into partitions
set.seed(3451)
inds <- partition(transformed_data$food_category, p = c(train = 0.7, valid = 0.2, test = 0.1))
#str(inds)
#> List of 3
#>  $ train: int [1:81] 2 3 6 7 8 10 11 18 19 20 ...
#>  $ valid: int [1:34] 1 12 14 15 27 34 36 38 42 48 ...
#>  $ test : int [1:35] 4 5 9 13 16 17 25 39 41 45 ...

train <- transformed_data[inds$train, ]
X_train=dplyr::select(train,-cost)
y_train=train$cost

valid <- transformed_data[inds$valid, ]
X_valid=dplyr::select(valid,-cost)
y_valid=valid$cost


test <- transformed_data[inds$test, ]
X_test=dplyr::select(test,-cost)
y_test=test$cost
```

Question 1{data-orientation=rows}
=====================================  
### What is the average income of the customers who shop at CFM?
Column {}
-------------------------------------
    
```{r}
avg_income_count <- CFM_data %>% dplyr::select(avg..yearly_income) %>% group_by(avg..yearly_income) %>% arrange(avg..yearly_income) 
order <- c("$10K - $30K","$30K - $50K","$50K - $70K","$70K - $90K","$90K - $110K","$110K - $130K","$130K - $150K","$150K +")

ggplot(avg_income_count,aes(avg..yearly_income)) + geom_bar() + theme(axis.text.x = element_text(angle = 0)) + ggtitle("Average Yearly Income")
```

Question 2 {data-orientation=rows}
=====================================   
### What is the Education distribution trend between the male and the female customers ?
  Column {}
-------------------------------------
```{r}
Ed_gender <- CFM_data %>% dplyr::select(gender,education) %>% arrange(gender) %>% group_by(gender,education)
#Ed_gender %>% count()
ggplot(Ed_gender, aes(x=education,fill=education)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) + facet_wrap(~gender) + ggtitle(("Gender distribution in \nEducation Qualification"))
``` 

 Question 3 {data-orientation=rows}
=====================================   
### Membership distribution in different income ranges?
  Column {}
-------------------------------------
```{r}
card_income <- CFM_data %>% dplyr::select(member_card,avg..yearly_income) %>% group_by(member_card,avg..yearly_income)
#card_income %>% count()
ggplot(card_income, aes(x=avg..yearly_income,fill=member_card)) + geom_bar()+ theme(axis.text.x = element_text(angle = 0)) + ggtitle("Memebership distribution \n in different income ranges")
```


Question 4 {data-orientation=rows}
=====================================  
### What are the top 3 promotional taglines that make the most sales for CFM?
Column {}
-------------------------------------
   

```{r}
promotion_data <- CFM_data %>% dplyr::select(promotion_name) %>% group_by(promotion_name)
promo_count<-promotion_data %>% group_by(promotion_name) 
#promo_count
ggplot(promo_count, aes(x=promotion_name),color=promotion_name) + geom_bar(fill = '#003366') + theme(axis.text.x = element_text(angle = 90))
```   

Question 5 {data-orientation=rows}
===================================== 
### Are there any brand preferences for customers with different occupations?
column {}
-------------------------------------
    
```{r fig.height=5, fig.width=15}

    brand_ocp <- CFM_data %>% dplyr::select(occupation,brand_name) %>% group_by(occupation)
    ggplot(brand_ocp, aes(x=occupation,fill=brand_name)) + geom_bar() + ggtitle("Band's favored by different Occupation")
```

Question 6 {data-orientation=rows}
===================================== 
### Which type of store is most profitable for CFM?
Column {}
-------------------------------------
```{r}
sales_type_rev <- tapply(CFM_data$store_sales.in.millions., CFM_data$store_type, FUN=sum)
#sales_type_rev
p <- barplot(tapply(CFM_data$store_sales.in.millions., CFM_data$store_type, FUN=sum),ylab="in Millions",main = "Store Sales in Differnt Store Types")
```

Question 7 {data-orientation=rows}
===================================== 
### Which state is most and least profitable for CFM?
Column {}
-------------------------------------
```{r}
sales_city <- CFM_data %>% dplyr::select(store_city,store_sales.in.millions.)
#sales_city
sales_city_rev <- tapply(CFM_data$store_sales.in.millions.,CFM_data$store_city,FUN=sum)
#sales_city_rev
barplot(sales_city_rev,las=2,ylab="in Millions",main="Revenue state-wise")
```

Question (8)a {data-orientation=rows}
===================================== 
### Predicting the media campaign cost for Convenient Food Mart (CFM).

column {}
-------------------------------------
```{r}
#XGBRegressor


dtrain <- xgb.DMatrix(data = as.matrix(X_train), label = y_train)
dtest <- xgb.DMatrix(data = as.matrix(X_test), label = y_test)
# Instantiate the model
model <- xgboost(data = dtrain, # the data   
                 nround = 350, # max number of boosting iterations)
                    max.depth = 5
                 )
```
column {}
-------------------------------------
```{r message=FALSE, warning=FALSE}
# Fit the model to the data
pred <- predict(model, dtest)
mse = mean((y_test - pred)^2)
mae = caret::MAE(y_test, pred)
rmse = caret::RMSE(y_test, pred)
cat("MSE: ", mse, "\nMAE: ", mae, "\nRMSE: ", rmse)
postResample(y_test,pred)

df = data.frame(y_test, pred)
df$id = 1:nrow(df)
```
Row {}
-------------------------------------
```{r}
ggplot() + geom_point(data = df, aes(x = id, y = y_test, color = 'y_test')) + 
  geom_point(data = df, aes(x=id, y = pred, color = 'pred',)) +
  theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = seq(0, 6000, by = 500))+
  ylab('Cost')
#confusion_mat = as.matrix(table(Actual_Values = y_test, Predicted_Values = pred)) 
#confusion_mat
importance_matrix = xgb.importance(colnames(dtrain), model = model)
xgb.plot.multi.trees(feature_names = names(CFM_data), 
                     model = model)
```
Question 8(b) {data-orientation=rows}
===================================== 
### Predicting the media campaign cost for Convenient Food Mart (CFM).
column {}
-------------------------------------
```{r message=FALSE, warning=FALSE}
#LGBMRegressor
library(lightgbm)
library(caret)
library(ggplot2) 
library(stats)

dtrain = lgb.Dataset(as.matrix(X_train), label = y_train)

dtest = lgb.Dataset.create.valid(dtrain,as.matrix(X_test), label = y_test)

#dvalid = lgb.Dataset.create.valid(dtrain,as.matrix(X_valid),label = y_valid)
set.seed(123)
valids = list(test = dtest)
model <- lgb.train(
  params = list(
    objective = "regression"
  , metric = "l2"
  , min_data = 1L
  , learning_rate = .3
  , nrounds = 10L
  ), 
  valids = list(test = dtest),
  data = dtrain
)


#model = lightgbm(boosting_type = 'gbdt', objective = "regression", metric = 'mae', dtrain, nrounds = 500) 
pred <- predict(model,dat = as.matrix(X_test))
```
column {}
-------------------------------------
```{r}
mse = mean((y_test - pred)^2)
mae = caret::MAE(y_test, pred)
rmse = caret::RMSE(y_test, pred)

cat("MSE: ", mse, "\nMAE: ", mae, "\nRMSE: ", rmse)
postResample(y_test,pred)

df = data.frame(y_test, pred)
df$id = 1:nrow(df)
options(repr.plot.width = 20, repr.plot.height =2)
```
Row {}
-------------------------------------
```{r}
ggplot() + geom_point(data = df, aes(x = id, y = y_test, color = 'y_test')) + 
  geom_point(data = df, aes(x=id, y = pred, color = 'pred')) + scale_x_continuous(breaks = seq(0, 6000, by = 500))+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab('Cost')

# feature importance
tree_imp = lgb.importance(model, percentage = TRUE)
lgb.plot.importance(tree_imp, top_n = 10L, measure = "Gain")
  #GAIN COVER Frequency 
```
   
 
