---
title: "Food Mart"
always_allow_html: true
output:
  pdf_document: default
  word_document: default
  html_document: 
    df_print: paged
date: "2022-10-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Loading Packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyr)
library(ggplot2)
library(dplyr)
#install.packages('corrplot')
library(corrplot)
library(xgboost)
library(caret)
library(splitTools)
library(ranger)
library(lightgbm)
library(caret)
library(ggplot2) 
library(stats)
library(caret)
library(randomForest)
library(ggplot2)
```
# About FoodMart:
Food Mart (CFM) is a chain of convenience stores in the United States. The private company's headquarters are located in Mentor, Ohio, and there are currently approximately 325 stores
located in the US. Convenient Food Mart operates on the franchise system.
Food Mart was the nation's third-largest chain of convenience stores as of 1988. The NASDAQ exchange dropped Convenient Food Mart the same year when the company failed to meet financial reporting requirements. Carden & Cherry advertised Convenient Food Mart with the Ernest character in the 1980s.



# Loadin Data
```{r}
#Load Data
CFM_data <- read.csv("C:/Users/bssup/Documents/fall22/R/project/Food_mart/data.csv")
```

# Correlation Matrix
This will show correlation between different variables.
Reasons of computing correlation matrix:
To summarize a large amount of data where the goal is to see patterns. In our example above,
the observable pattern is that all the variables highly correlate with each other. 

```{r}
#Correlation Matrix
store_data <- dplyr::select(CFM_data,c('store_sales.in.millions.','store_cost.in.millions.','unit_sales.in.millions.','total_children','avg_cars_at.home.approx.','num_children_at_home','avg_cars_at.home.approx..1','SRP','gross_weight','net_weight','recyclable_package','low_fat','units_per_case','store_sqft','grocery_sqft','frozen_sqft','meat_sqft','coffee_bar','video_store','salad_bar','prepared_food','florist','cost'))
#store_data
cor_data <- cor(store_data)
#head(cor_data)
corrplot(cor_data,method="shade",tl.col = "black",title = "\n\n Correlation Plot Of Store Data",tl.srt=65,cl.length = 5,cl.align="l",tl.cex=0.95,  number.cex=0.8,col = colorRampPalette(c("yellow","lightblue","navyblue"))(100))

```

```{r message=FALSE, warning=FALSE}

library(ggplot2)
library(reshape2)
data <- CFM_data
library(plotly)

fig <- plot_ly(data, x = ~education, color = ~gender, type = 'histogram', 
               title = 'To visualize the education level of people', 
               text = ~education) %>%
  layout(title = 'Sampled Results', 
         bargap = 0.2,
         bargroupgap = 0.1,
         plot_bgcolor = 'pink')

fig
```
`   qa    
# Count Plot Promotion Name
```{r}
promotion_data <- CFM_data %>% dplyr::select(promotion_name) %>% group_by(promotion_name)
promo_count<-promotion_data %>% group_by(promotion_name) 
ggplot(promo_count, aes(x=promotion_name),color=promotion_name) + geom_bar(fill = '#003366') + theme(axis.text.x = element_text(angle = 90))
```

# Pie Chat to Visualize the food_category in dataset

```{r}
food_cat <- CFM_data %>% dplyr::select(food_category) %>% group_by(food_category)%>% count()
food_cat<- arrange(food_cat)
piepercent<- round(100*food_cat$n/sum(food_cat$n), 1)
#pie(food_cat$n,piepercent)



pie(food_cat$n, labels = food_cat$food_category, main = "Food Category pie chart",col = rainbow(length(food_cat$n)), radius = 1, cex = 0.6) 
```

# Observation:
From the above pie chart most used food_category such as 
1.Vegetables 
2.Snack Foods
3.Dairy these products we used in our daily life


# To Visualize the promotion_names
```{r}
promotion_data <- CFM_data %>% dplyr::select(promotion_name) %>% group_by(promotion_name)
promo_count<-promotion_data %>% group_by(promotion_name) 
#promo_count
ggplot(promo_count, aes(x=promotion_name),color=promotion_name) + geom_bar(fill = '#003366') + theme(axis.text.x = element_text(angle = 90))

```
# Observation
From the above graph we found out that the most sales promotors names are:
1. Weekend Markdown
2. Two Day sales
3. Price Slashers

# Visualize the Average yearly Income 

```{r}
avg_income_count <- CFM_data %>% dplyr::select(avg..yearly_income) %>% group_by(avg..yearly_income) %>% arrange(avg..yearly_income) 
order <- c("$10K - $30K","$30K - $50K","$50K - $70K","$70K - $90K","$90K - $110K","$110K - $130K","$130K - $150K","$150K +")
#avg_income_count
ggplot(avg_income_count,aes(avg..yearly_income)) + geom_bar() + theme(axis.text.x = element_text(angle = 0)) + ggtitle("Average Yearly Income")

```

# Observation 
We found out that the average yearly income of the customers is $30k - $50k.

# To Visualize the Education with gender in the dataset 
```{r}
Ed_gender <- CFM_data %>% dplyr::select(gender,education) %>% arrange(gender) %>% group_by(gender,education)
#Ed_gender %>% count()
ggplot(Ed_gender, aes(x=education,fill=education)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) + facet_wrap(~gender) + ggtitle(("Gender distribution in \nEducation Qualification"))
```

# To Visualize occupation and brand_name divided among gender
```{r fig.height=5, fig.width=15}
  brand_ocp <- CFM_data %>% dplyr::select(occupation,brand_name) %>% group_by(occupation)
    ggplot(brand_ocp, aes(x=occupation,fill=brand_name)) + geom_bar() + ggtitle("Band's favored by different Occupation")

```

# Observation

For Male:
1. Most male have the partial high school degree.
2. Females with graduate degree are low in number.

# Visualize the yearly income with member card
Member card is the membership card that the customer owns.
```{r}
card_income <- CFM_data %>% dplyr::select(member_card,avg..yearly_income) %>% group_by(member_card,avg..yearly_income)
#card_income %>% count()
ggplot(card_income, aes(x=avg..yearly_income,fill=member_card)) + geom_bar()+ theme(axis.text.x = element_text(angle = 0)) + ggtitle("Memebership distribution in different income ranges")
```

# Visualize the education with occupation
```{r}
ed_oc <- CFM_data %>% dplyr::select(occupation,education) %>% group_by(occupation,education)
#ed_oc %>% count()
ggplot(ed_oc, aes(x=occupation,fill=education)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) + scale_fill_hue(limits = c("Partial High School","High School Degree","Partial College","Bachelors Degree","Graduate Degree")) + ggtitle("occupation Vs Education")

```

# Visualize the average income with occupation
```{r}
Income_ocup <- CFM_data %>% dplyr::select(avg..yearly_income,occupation) %>% group_by(occupation,avg..yearly_income)
#Income_ocup %>% count()
ggplot(Income_ocup,aes(x=avg..yearly_income,fill=occupation)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) + ggtitle("Income Distribution and Occupation")

```

# Observation 

1. Professional employee average earns around $50k +. 
2. Skilled manual employee and manual employee earns $30k +.
3. Management employee earns $30k +.
4. Clerical employee earns $30k +.

# Visualize the average yearly income with occupation
```{r}
Income_ocup <- CFM_data %>% dplyr::select(avg..yearly_income,occupation) %>% group_by(occupation,avg..yearly_income)
#Income_ocup %>% count()
ggplot(Income_ocup,aes(x=avg..yearly_income,fill=occupation)) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) + ggtitle("Income Distribution and Occupation")

```

# Visualise store sales with the state the store is located.

```{r}
City_income <- CFM_data %>% dplyr::select(store_city,store_sales.in.millions.) %>% group_by(store_city,store_sales.in.millions.)
ggplot(City_income, aes(x=store_city),y=store_sales.in.millions.) + geom_bar()+ theme(axis.text.x = element_text(angle = 90))

```


# Visualize store type and the state it is located
```{r}
sales_type_rev <- tapply(CFM_data$store_sales.in.millions., CFM_data$store_type, FUN=sum)
#sales_type_rev
p <- barplot(tapply(CFM_data$store_sales.in.millions., CFM_data$store_type, FUN=sum),ylab="in Millions",main = "Store Sales in Differnt Store Types")

```

# Visualize store sales in each state
```{r}
sales_city <- CFM_data %>% dplyr::select(store_city,store_sales.in.millions.)
#sales_city
sales_city_rev <- tapply(CFM_data$store_sales.in.millions.,CFM_data$store_city,FUN=sum)
#sales_city_rev
barplot(sales_city_rev,las=2,ylab="in Millions",xlab="city",main="Revenue state-wise")
```


```{r message=FALSE, warning=FALSE}
library(CatEncoders)
transformed_data = CFM_data
#define original categorical labels and convert all the data for modeling 
label = LabelEncoder.fit(CFM_data$food_category)
transformed_data$food_category = transform(label, CFM_data$food_category)

label = LabelEncoder.fit(CFM_data$food_department)
transformed_data$food_department = transform(label, CFM_data$food_department)

label = LabelEncoder.fit(CFM_data$food_family)
transformed_data$food_family = transform(label, CFM_data$food_family)

label = LabelEncoder.fit(CFM_data$promotion_name)
transformed_data$promotion_name = transform(label, CFM_data$promotion_name)

label = LabelEncoder.fit(CFM_data$marital_status)
transformed_data$marital_status = transform(label, CFM_data$marital_status)

label = LabelEncoder.fit(CFM_data$gender)
transformed_data$gender = transform(label, CFM_data$gender)

label = LabelEncoder.fit(CFM_data$education)
transformed_data$education = transform(label, CFM_data$education)

label = LabelEncoder.fit(CFM_data$member_card)
transformed_data$member_card = transform(label, CFM_data$member_card)

label = LabelEncoder.fit(CFM_data$occupation)
transformed_data$occupation = transform(label, CFM_data$occupation)

label = LabelEncoder.fit(CFM_data$avg..yearly_income)
transformed_data$avg..yearly_income = transform(label, CFM_data$avg..yearly_income)

label = LabelEncoder.fit(CFM_data$sales_country)
transformed_data$sales_country = transform(label, CFM_data$sales_country)

label = LabelEncoder.fit(CFM_data$brand_name)
transformed_data$brand_name = transform(label, CFM_data$brand_name)

label = LabelEncoder.fit(CFM_data$houseowner)
transformed_data$houseowner = transform(label, CFM_data$houseowner)

label = LabelEncoder.fit(CFM_data$store_type)
transformed_data$store_type = transform(label, CFM_data$store_type)

label = LabelEncoder.fit(CFM_data$store_city)
transformed_data$store_city = transform(label, CFM_data$store_city)

label = LabelEncoder.fit(CFM_data$store_state)
transformed_data$store_state = transform(label, CFM_data$store_state)

label = LabelEncoder.fit(CFM_data$media_type)
transformed_data$media_type = transform(label, CFM_data$media_type)

#(transformed_data)
```
```{r}
#Correlation Matrix
store_data <- dplyr::select(transformed_data,c('store_sales.in.millions.','store_cost.in.millions.','unit_sales.in.millions.','total_children','avg_cars_at.home.approx.','num_children_at_home','avg_cars_at.home.approx..1','SRP','gross_weight','net_weight','recyclable_package','low_fat','units_per_case','store_sqft','grocery_sqft','frozen_sqft','meat_sqft','coffee_bar','video_store','salad_bar','prepared_food','florist','promotion_name','cost'))
#store_data
cor_data <- cor(store_data)
#head(cor_data)
corrplot(cor_data,method="shade",tl.col = "black",title = "\n\n Correlation Plot Of Store Data",tl.srt=65,cl.length = 5,cl.align="l",tl.cex=0.95,  number.cex=0.8,col = colorRampPalette(c("yellow","lightblue","navyblue"))(100))

```

```{r}

#split the data 


# Split data into partitions
set.seed(3451)
inds <- partition(transformed_data$food_category, p = c(train = 0.7, valid = 0.2, test = 0.1))
#str(inds)
#> List of 3
#>  $ train: int [1:81] 2 3 6 7 8 10 11 18 19 20 ...
#>  $ valid: int [1:34] 1 12 14 15 27 34 36 38 42 48 ...
#>  $ test : int [1:35] 4 5 9 13 16 17 25 39 41 45 ...

train <- transformed_data[inds$train, ]
X_train=dplyr::select(train,-cost)
y_train=train$cost

valid <- transformed_data[inds$valid, ]
X_valid=dplyr::select(valid,-cost)
y_valid=valid$cost


test <- transformed_data[inds$test, ]
X_test=dplyr::select(test,-cost)
y_test=test$cost

#print(X_train)
#print(y_train)
#dim(X_train)
#length(y_train)
#dim(X_test)
#length(y_test)
#dim(X_valid)
#length(y_valid)
```

```{r}
##Explore the data.
library(MASS)
library(ggplot2)
ggplot(X_train, aes(y_train)) + geom_density(fill="blue")
ggplot(X_train, aes(log(y_train))) + geom_density(fill="blue")
ggplot(X_train, aes(sqrt(y_train))) + geom_density(fill="blue")
```

```{r}
model1 = lm(log(y_train)~., data=X_train)
summary(model1)
par(mfrow=c(2,2))
plot(model1)
```

```{r}
#Random forest 
#library(randomForest)
#rf.fit <- randomForest(y_train ~ ., data=X_train, ntree=1000,
#                       keep.forest=FALSE, importance=TRUE)
#rf.fit
#pred <- rf.predict( dtest)
```


```{r message=FALSE, warning=FALSE}
#XGBRegressor
dtrain <- xgb.DMatrix(data = as.matrix(X_train), label = y_train)
dtest <- xgb.DMatrix(data = as.matrix(X_test), label = y_test)
# Instantiate the model
model <- xgboost(data = dtrain, # the data   
                 nround = 350, # max number of boosting iterations)
                    max.depth = 5
                 )
# Fit the model to the data
pred <- predict(model, dtest)
mse = mean((y_test - pred)^2)
mae = caret::MAE(y_test, pred)
rmse = caret::RMSE(y_test, pred)
postResample(y_test,pred)
cat("MSE: ", mse, "\nMAE: ", mae, "\nRMSE: ", rmse)
df = data.frame(y_test, pred)
df$id = 1:nrow(df)
ggplot() + geom_point(data = df, aes(x = id, y = y_test, color = 'y_test')) + 
  geom_point(data = df, aes(x=id, y = pred, color = 'pred',)) +
  theme(plot.title = element_text(hjust = 0.5)) + scale_x_continuous(breaks = seq(0, 6000, by = 400))+
  ylab('Cost')
importance_matrix = xgb.importance(colnames(dtrain), model = model)
importance_matrix
xgb.plot.multi.trees(feature_names = names(CFM_data), 
                     model = model)
postResample(y_test,pred)

```

```{r message=FALSE, warning=FALSE}
#LGBMRegressor

dtrain = lgb.Dataset(as.matrix(X_train), label = y_train)

dtest = lgb.Dataset.create.valid(dtrain,as.matrix(X_test), label = y_test)

#dvalid = lgb.Dataset.create.valid(dtrain,as.matrix(X_valid),label = y_valid)
set.seed(123)
valids = list(test = dtest)
model <- lgb.train(
  params = list(
    objective = "regression"
  , metric = "l2"
  , min_data = 1L
  , learning_rate = .3
  , nrounds = 10L
  ), 
  valids = list(test = dtest),
  data = dtrain
)
#model = lightgbm(boosting_type = 'gbdt', objective = "regression", metric = 'mae', dtrain, nrounds = 500) 
pred <- predict(model,dat = as.matrix(X_test))
length(pred)
length(y_test)
mse = mean((y_test - pred)^2)
mae = caret::MAE(y_test, pred)
rmse = caret::RMSE(y_test, pred)
cat("MSE: ", mse, "\nMAE: ", mae, "\nRMSE: ", rmse)
postResample(y_test,pred)

df = data.frame(y_test, pred)
df$id = 1:nrow(df)
options(repr.plot.width = 20, repr.plot.height =2)
ggplot() + geom_point(data = df, aes(x = id, y = y_test, color = 'y_test')) + 
  geom_point(data = df, aes(x=id, y = pred, color = 'pred')) + scale_x_continuous(breaks = seq(0, 6000, by = 400))+
  theme(plot.title = element_text(hjust = 0.5)) +
  ylab('Cost')

# feature importance
tree_imp = lgb.importance(model, percentage = TRUE)
lgb.plot.importance(tree_imp, top_n = 10L, measure = "Gain")

```